---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {diades.atlas}

<!-- badges: start -->
[![R build status](https://github.com/inrae/diades.atlas/workflows/R-CMD-check-docker-renv/badge.svg)](https://github.com/inrae/diades.atlas/actions)
[![Codecov test coverage](https://codecov.io/gh/inrae/diades.atlas/branch/main/graph/badge.svg)](https://codecov.io/gh/inrae/diades.atlas?branch=main)
<!-- badges: end -->

A Shiny application to explore data.

# Installation

```{r, eval=FALSE}
remotes::install_github('inrae/diades.atlas')
```

# Websites for validations :

- Package documentation : https://inrae.github.io/diades.atlas/ 
- Package documentation (development version) : https://inrae.github.io/diades.atlas/dev/ 
- Shiny App : https://connect.thinkr.fr/diadesatlasui18n/

# Translation

Toutes les traductions sont stockées dans des fichiers csv dans "inst/". S'il y a des corrections à apporter, des textes existants sur l'app à traduire, seuls ces 3 fichiers sont à modifier: 

+ `inst/translation_help.csv`
+ `inst/translation_iucn.csv`
+ `inst/translation.csv`

Après modifications dans une branche, proposez une Pull Request vers la branche _main_:  

- Sur la page d'acceuil du projet, cliquer sur Pull Request
- Puis sur "New Pull Request"
- Dans la partie "Compare Changes", choisir votre branche dans le menu déroulant de droite "compare:"
- Puis "Create Pull Request", 2 fois.
- Vous pouvez vérifier dans l'onglet 'Files changed' ce que vous avez modifié.

Une fois la PR acceptée, les modifications devraient apparaître dans l'app au prochain déploiement. Il n'y a rien de puls à faire.

Pour plus d'informations, ou pour les instructions pour préparer l'application à recevoir de nouvelles langues ou zones à traduire, se réferrer au fichier "dev/translation.Rmd".  

Dans tous les cas, les développeurs devront 
- avoir accès à la base de données PostGis (Voir "Backend requirement" ci-dessous)
- suivre le contenu de "dev/translation.Rmd" lors de chaque mises à jour.

# Retrieve package and develop

You can get this package and propose modifications of the code.

1. Clone in RStudio
  + File > New Project > Version Control > Git
  + Use URL: https://github.com/inrae/diades.atlas
  
2. In the RStudio "Git" Pane
  + Create a "New branch" with your name
  
3. Define your credentials for the database
  + Create a ".Renviron" file with `usethis::edit_r_environ()`
  + Use the correct names and passwords to fill as below
  + This file will stay on your computer
```
POSTGRES_DBNAME="diades"
POSTGRES_HOST="host-url"
POSTGRES_USER="username"
POSTGRES_PORT=5432
POSTGRES_PASS="user-password"
```
  + Restart RStudio
  + Verify these variables are correctly read
```r
Sys.getenv("POSTGRES_USER")
```

4. You may need to install dependencies with the same packages than the developers

```r
renv::restore()
```

_{renv} may be tricky with multiple OS. Be kind._

5. Run `pkgload::load_all()` in the R Console
  + This loads all existing functions stored in "R/" directory
  
6. Open a Rmd in "data-raw/" like "data-raw/be-page4-future.Rmd"
  + You should be able to run it chunk by chunk
  + Do not forget to close connection to the database at the end

7. You can test modifications of the code in "R/"
  + After each modification, run `pkgload::load_all()` before testing in the Rmd

# Exploration of database out of Shiny application

You can use this package without opening the Shiny application.  
See vignette "00-exploration-of-data"

# Put in Production

Update Docker for deployment: https://gitlab.irstea.fr/diades/diades.atlas.deploy/

On https://gitlab.irstea.fr/diades/diades.atlas.deploy/, there are also details on how to test the docker-compose locally. At least try to build it locally before release.

```sh
docker build -t gitlab-registry.irstea.fr/diades/diades.atlas.deploy/app .
```

# Backend requirement 

## PostGis

To update the translations or to run the shiny application locally, the PostGis database is required.

To be able to connect to a PostGis instance the user needs to define some golem config parameters and environment variables.

### golem config parameters

Those are stored in `inst/golem-config.yml` and can be modified as such:

```{r eval=FALSE}
# database name
golem::amend_golem_config(
  "POSTGRES_DBNAME",
  "diades" # Changer ici le nom de la base de données
)
# database host URL
golem::amend_golem_config(
  "POSTGRES_HOST",
  "localhost"  # Changer ici l'URL de la base
)
# database port access
golem::amend_golem_config(
  "POSTGRES_PORT",
  "5432" # Changer ici le port
)
```

**If you change these lines to set the real values, be sure to set them back with the examples below before commit**

### Environment variables

The POSTGRES_USER and POSTGRES_PASS __must not be versioned__ with the app.

To avoid setting those environment variables every time you want to run the app,
they should be saved in a local .Renviron file. To define one for your own copy of
the project run this command:

```{r eval=FALSE}
usethis::edit_r_environ(scope = "project")
```

and add the following lines

```{r eval=FALSE}
# database username with Read access
POSTGRES_USER=diadesatlas_owner # Changer ici le username
# database password for the user
POSTGRES_PASS=thinkrpassword # Changer ici le password
```

Again make sure not to track this project .Renviron with git.

### Dev - ThinkR PostGis DB

If you do not have direct access to the database, a copy can be used for development.
Please refer to the doc in the {diadesdata} repo on ThinkR's forge (restricted access): Section "Pull and Use".

### Checks

Verify connexion to database works  
_Whether you are connected to the development or the production database, this should work correctly_

```{r, eval=FALSE}
pkgload::load_all()
session <- new.env()
connect(session)
con <- get_con(session)

DBI::dbListTables(con)
DBI::dbDisconnect(con)
```

## Dev - Mongo

To run the Shiny application locally, you need a Mongo database. For dev, you need to run the following on your machine:

```
docker run --name=mongo --rm -p 2811:27017 -e MONGO_INITDB_ROOT_USERNAME=Colin -e MONGO_INITDB_ROOT_PASSWORD=AsAboveSoBelow789123 mongo:4.0
```

Stop it at the end of your development process

```
docker kill mongo
```

# Dev - Data update requires update unit test expected outputs

If the database is updated, model outputs may change. 
It may be necessary to run simulation before unit tests.

```{r, eval=FALSE}
source(here::here("data-raw/altas_simulation.R"))
```

# Dev - update and include vignettes in the package

Vignettes are compiled manually by developers. 
Raw vignettes are stored in "data-raw". 
Instructions to compile them are in the vignette itself.  

However, they can all be prepared from this script: 

```{r, eval=FALSE}
remotes::install_local(upgrade = "never", force = TRUE)

file.copy(here::here("dev", "translation.Rmd"),
          here::here("data-raw", "translation.Rmd"))

all_vignettes <- c(
  "aa-a-exploration_data",
  "aa-data-exploration-and-preparation",
  "bb-page1-catch-bycatch",
  "bc-page2-present",
  "bd-page3-climate-change",
  "be-page4-future", 
  "translation")

for (vignette_name in all_vignettes) {
  # vignette_name <- all_vignettes[6]
  vignette_file <- paste0(vignette_name, ".Rmd")
  
  rmarkdown::render(
    input = here::here(file.path("data-raw", vignette_file)),
    output_format = "rmarkdown::html_vignette",
    output_options = list(toc = TRUE),
    output_file = here::here(file.path("vignettes", vignette_file))
  )
  
  # Add header for title
  lines <- readLines(here::here(file.path("vignettes", vignette_file)))
  
  cat(
    glue::glue('---
title: ".{vignette_name}."
output: rmarkdown::html_vignette
vignette: >
  %\\VignetteIndexEntry{.{vignette_name}.}
  %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}
---
', .open = ".{", .close = "}."),
lines,
sep = "\n", 
file = here::here(file.path("vignettes", vignette_file))
  )
}
file.remove(here::here("data-raw", "translation.Rmd"))
```



