---
title: "00-exploration-of-data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{00-exploration-of-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Installation

Install the present package and use its core functions

```{r, eval=FALSE}
remotes::install_github('inrae/diades.atlas')
```

## Packages
```{r}
library(dplyr)
library(DBI)
library(ggplot2)
library(RPostgres)
library(diades.atlas)
# pkgload::load_all()
```

**You can use this package without opening the Shiny application**

## Connect to the database

```{r, eval=FALSE}
se <- new.env()

conn_eurodiad <- dbConnect(
  RPostgres::Postgres(), 
  dbname = 'eurodiad',
  host = '<your-host>',
  port = 5432,
  user = '<your-username>',
  password = rstudioapi::askForPassword("Database password"))

session$userData$con <-conn_eurodiad
```

```{r}
# For ThinkR only
se <- new.env()
conn_eurodiad <- connect(session = se)
DBI::dbListTables(conn_eurodiad)
```

## Explore data

It is recommended to use `tbl()` instead of `dbGetQuery()` to connect to a table in the database.  
`tbl()` does not download the table, it only connect to the database and let it run the query. This means that if you run a `filter()` or a `mutate()` after that, the database executes it, not R, which is more efficient.  
However, at some point, some of your {tidyverse} operations can not be realised or translated in SQL, hence, you need to download the resulting dataset using `collect()`. Then you can continue as usual.

For instance. 

- `dbGetQuery()` download data in R
  - `filter()` is executed by R in your R session
- Note that `!!` is a specific to using {dplyr} with SQL there.
  + This is because the variable after it (e.g. `species_id`) is defined in the R session, but not in the database. Therefore, before sending the SQL query to the database, R has to transform the variable by its real value in R. Otherwise, it will send the word `"species_id"` which does not make sense for the SQL database, instead of the number you wanted to put.


```{r}
# Create a function
get_data_dbi <- function(conn_eurodiad, species_id, scenario) {
  query <- "SELECT * FROM diadesatlas.v_hybrid_model"
  
  dbGetQuery(conn_eurodiad, query) %>%
    filter(species_id %in% !!species_id,
           climatic_scenario %in% !!scenario) 
}

# Use it
get_data_dbi(conn_eurodiad,
             species_id = c(6),
             scenario = 'rcp85')
```

- `tbl()` only connects to the table, only a glimpse of the data is presented
  - `filter()` is run by the SQL database

```{r}
get_data_tbl <- function(conn_eurodiad, species_id, scenario) {
  tbl(conn_eurodiad, "v_hybrid_model") %>% 
    filter(species_id %in% !!species_id,
           climatic_scenario %in% !!scenario) 
}

# Use it
get_data_tbl(conn_eurodiad,
             species_id = c(6),
             scenario = 'rcp85')
```

- If you need to create some {ggplot2} figures after that, you can collect

```{r}
get_data_tbl_collect <- function(conn_eurodiad, species_id, scenario) {
  tbl(conn_eurodiad, "v_hybrid_model") %>% 
    filter(species_id %in% !!species_id,
           climatic_scenario %in% !!scenario) %>% 
    # From collect(), the full data is in the R session
    collect()
}

# Use it
get_data_tbl_collect(conn_eurodiad,
                     species_id = c(6),
                     scenario = 'rcp85') %>% 
  filter(basin_id == 1) %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = hsi))
```

- Note that you can use a full SQL query with `tbl()`

```{r}
get_data_tbl_query_collect <- function(conn_eurodiad, species_id, scenario) {
  query <- "SELECT * FROM diadesatlas.v_hybrid_model"
  
  tbl(conn_eurodiad, sql(query)) %>% 
    # Adds a filter to be run as SQL in the database
    filter(species_id %in% !!species_id,
           climatic_scenario %in% !!scenario) %>% 
    # From collect(), the full data is in the R session
    collect()
}

# Use it
get_data_tbl_query_collect(conn_eurodiad,
                           species_id = c(6),
                           scenario = 'rcp85') %>% 
  filter(basin_id == 1) %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = hsi))
```


## Disconnect from the database

```{r}
DBI::dbDisconnect(conn_eurodiad)
```

