---
title: "aa-partie-exploration-des-donnees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aa-partie-exploration-des-donnees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette explores the tables in the database and ensures that R has access to the tables.
It also transforms the data in json.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE # Set to TRUE when data OK for test, but not on GitHub
)

```

```{r setup}
library(diades.atlas)
# pkgload::load_all(here::here())
# renv::install("dm")
# library(dm)
library(geojson)
library(dplyr)
library(leaflet)
library(ggplot2)
library(sf)
```

## Data exploration

### Connect to database

Do not forget to set environment variables in .Renviron
```{r}
session <- shiny::MockShinySession$new() #new.env()
connect(session)
con <- get_con(session)
```

```{r, eval=FALSE}
# Get the names of all the tables in the schema
all_schemas <- DBI::dbGetQuery(con, "SELECT table_name FROM information_schema.tables WHERE table_schema='diadesatlas'")
all_schemas

# Display the head of all
for (i in all_schemas$table_name) {
  print(dplyr::tbl(con, dbplyr::in_schema("diadesatlas", i)))
}

casestudy <- tbl(con, dbplyr::in_schema("diadesatlas", "casestudy")) %>% 
  filter(publishable == TRUE)
casestudy
```

### Function to collect data
```{r, eval=FALSE}
# empty
# dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "cices51")) %>% collect()
# services ?
# dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "cices")) %>% collect()

# geometry of all basin
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "basin_location")) %>% collect()
# Model results over the years
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "hybrid_model_result")) %>% collect()
# Abundance of species in a given basin over the years
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "abundance")) %>% collect()
# ?
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "climatic_model")) %>% collect()
# List of all basins
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "basin")) %>% collect()

# List of basin for case_study
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "casestudy")) %>% 
  filter(publishable == TRUE) %>% collect()
# List of basins id for case_study
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "casestudy_basin")) %>% collect()
# List of ecosystem services
# dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "ecosystem_service")) %>% collect()
# dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "ecosystem_service_new")) %>% collect()
# List of species
dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "species")) %>% collect()
```

## Prepare data for Home page
### Prepare json files
Create all json data used on the home page of the app
```{r, eval=FALSE}
# # List of case studies
# casestudy <- dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "casestudy")) %>% collect()
# # List of the corresponding basins
# casestudy_basin <- dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "casestudy_basin")) %>% collect()
# # Put them two together
# cstd <- full_join(casestudy, casestudy_basin) %>% filter(publishable) %>%  collect()


casestudy_new <- sf::st_read(con, 
                             query = "SELECT casestudy_id, casestudy_name, geom FROM diadesatlas.casestudy WHERE publishable = TRUE", geom = "geom") %>% 
  # fake basin_id to be compatabible with mod_b_second.R
  mutate(basin_id = 1:n()) %>% 
  st_transform(crs = 4326)

casestudy_new %>%
  geojson::as.geojson() %>%
  geojson::geo_write(here::here("inst/casestudy.json"))


# Get the features for drawing the polygons on the map
# qry <- "SELECT ST_Transform(diadesatlas.basin_outlet.simplified_geom, 4326) as geom, basin_id as basin_id FROM diadesatlas.basin_outlet"

# qry <- "SELECT diadesatlas.basin_outlet.simplified_geom as geom, basin_id as basin_id FROM diadesatlas.basin_outlet"
# 
# pols <- sf::st_read(con, query = qry, geom = "geom") %>% 
#   st_transform(crs = 4326) 
# 
# # i then read as an sf object
# pols %>%
#   filter(basin_id %in% casestudy_basin$basin_id) %>%
#   left_join(cstd)
# pols %>%
#   geojson::as.geojson() %>%
#   geojson::geo_write(here::here("inst/casestudy.json"))

# List of ecosystem services
# ecosystems <- dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "ecosystem_service")) %>% collect()
# ecosystems %>%
#   filter(presence == 1) %>%
#   jsonlite::write_json(here::here("inst/ecosystems.json"))
# Services
services <- dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "v_ecosystemic_services")) %>% collect()
services %>%
  distinct(category_id, subcategory_id, category_name, subcategory_name) %>%
  mutate(
    subcategory_id = gsub("99", "0", subcategory_id)
  ) %>%
  arrange(
    category_id, subcategory_id
  ) %>%
  mutate(
    subcategory_id = gsub("0", "99", subcategory_id)
  ) %>%
  jsonlite::write_json(here::here("inst/services.json"))

# eco_and_services <- full_join(ecosystems, services)

# Species
species <- dplyr::tbl(con, dbplyr::in_schema("diadesatlas", "species")) %>% 
  filter(active == TRUE) %>%
  collect()
# species %>%
#   filter(species_id %in% ecosystems$species_id) %>%
#   jsonlite::write_json(here::here("inst/species.json"))
```

### Explore ecosystem services table
```{r}
r <- list(lg = 'en')
datasets <- generate_datasets(con = con)
.globals <- getFromNamespace('.globals', 'shiny')

# simulate a species_list
session$options$golem_options$species_list <- datasets$species_list
.globals$domain <- session

golem::get_golem_options("species_list")

# debugonce(ecosystem_table)
ecosystem_table(species = "Alosa alosa", 
                case_study = 5, 
                ecosystem = '1_4',
                r = r,
                session = session) %>% 
DT::datatable()
```


### Stop connection

```{r}
DBI::dbDisconnect(con)
```



